---
title: "TIME-SERIES ANALYSIS"
author: "MAHALAKSHMI"
output: 
  html_document: 
    keep_md: yes
---

```{r global_options, include=FALSE,cache=FALSE}
knitr::opts_chunk$set(echo = FALSE,echo=FALSE, message=FALSE, warning=FALSE,error=TRUE)
```

Loading the relevant R packages

```{r packages}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggvis)
```

Loading the relevant datasets

```{r LOADING}
Mat1 <- read.table("C:/Users/Mash/Desktop/Kaggle/Time_Series_R/Mat1.csv",header=T,sep = ",",row.names =1)
Mat2 <- read.table("C:/Users/Mash/Desktop/Kaggle/Time_Series_R/Mat2.csv",header=T,sep = ",", row.names = 1)
```

Converting data to Time series format

```{r TS type}
Mat1 <- ts(Mat1$VALUE, start=c(1948,1), freq=12)
Mat2 <- ts(Mat2$VALUE, start=c(1948,1), freq=12)
str(Mat1)
str(Mat2)
```
Exploring the summary statistics

```{r SUMMARY}
summary(Mat1)
summary(Mat2)
```

Plotting the Time series behavior of the two data sets Mat1 and Mat2 with the regression line showing the general trend.The Mat1 data tends to be a lot more stationary showing no trend but MAt2 data shows an upward trend.

```{r PLOTTING TIME SERIES}
ts.plot(Mat1,Mat2,col=c("#20C345","red"))
abline(reg = lm(Mat1~time(Mat1)),col = c("#20C345"))
abline(reg = lm(Mat2~time(Mat2)),col = c("red"))
```

Showing the Decomposition plot

```{r using GGPLOT}
library(ggfortify)
#autoplot(Mat1,ts.colour = 'red')
#autoplot(Mat2,ts.colour = 'green')
autoplot(stl(Mat1, s.window = 'periodic'), ts.colour = 'green')
autoplot(stl(Mat2, s.window = 'periodic'), ts.colour = 'red')
```

Plotting the ACF for both the data.
Looks like a gradual slowdown for both the data sets indicating a AR model

```{r ACF}
autoplot(acf(Mat1,plot = FALSE))
autoplot(acf(Mat2,plot = FALSE))

```

Taking the Differentiation and then the ACF shows a gradual slowdown.

```{r DIFF}
autoplot(acf(diff(Mat1,plot = FALSE)))
autoplot(acf(diff(Mat2,plot = FALSE)))
```

Taking the PACF to see if there is a cut off or a gradual slowdown.
The Plot shows a Gradual slowdown indicating a ARIMA model

```{r PACF}
autoplot(pacf(diff(Mat1),plot = FALSE))

```
```{r ARIMA}
library(forecast)
Mat1.arima <-arima(Mat1, order = c(2,1,2))
Mat2.arima <- arima(Mat2,order = c(2,1,2))
Mat1.ar <-arima(Mat1, order = c(2,0,0))
Mat1.ma <-arima(Mat1, order = c(0,0,2))
print (Mat1.arima)
```


Plotting the details of the ARIMA (2,1,2) model

```{r PLOTTING}
plot(forecast(Mat1.arima, h=24))
plot(forecast(Mat2.arima, h=24))
```

